Continuous gravitational waves in the lab: recovering audio signals with a table-top~optical microphone

### 2021-10-08 paper/ifo-futureWork.tex
minor edits

Returning to the demonstration of gravitational wave analysis, single interferometers cannot yield directional information for signals; a large proportion of the directional information in gravitational-wave detections comes from the time delay offset between the signal\han{note} being recorded at two or more detectors.\cite{GW150914}
We could therefore extend this analysis to include data from two interferometers to extract directional information.
This would require increasing the sensitivity of the optical microphone to pick up the signal from a distant source instead of from a speaker attached directly to one of the mirrors of the interferometer.
Another extension would be to demonstrate the Doppler effect of the Earth's motion around the Sun, which needs to be considered in continuous-wave searches (see Section~\ref{sec:realCWSearches} and Ref.~\cite{JKS:1998}). 
One approach could be to modify the input audio signal to simulate Doppler modulation. 
% did we cut the turn-table idea? that was fun! - james
% yeah afraid so - it was nice, but the editor did not like it! :( - Hannah
% haha, okay. - james

>>>>

Returning to the demonstration of gravitational wave analysis, single interferometers cannot yield directional information for signals; a large proportion of the directional information in gravitational-wave detections comes from the time delay offset between the observation being recorded at two or more detectors.\cite{GW150914}
We could therefore extend this analysis to include data from two interferometers to extract directional information.
This would require increasing the sensitivity of the optical microphone to pick up the signal from a distant source instead of from a speaker attached directly to one of the mirrors of the interferometer.
Another extension would be to demonstrate the Doppler effect of the Earth's motion around the Sun, which needs to be considered in continuous-wave searches (see Section~\ref{sec:realCWSearches} and Ref.~\cite{JKS:1998}). 
One approach could be to modify the input audio signal to simulate Doppler modulation. 
% did we cut the turn-table idea? that was fun! - james
% yeah afraid so - it was nice, but the editor did not like it! :( - Hannah
% haha, okay. - james

### 2021-10-08 paper/ifo-wanderingTone.tex
grammar and spelling
minor edits

A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars, or variable accretion from a stellar companion for neutron stars in binaries such as LMXBs (see Ref.~\cite{SuvorovaEtAl:2017} and references therein).
Although continuous gravitational waves are nearly monochromatic, the long observing times ($\sim 1\,{\rm yr}$) mean that searches are sensitive to very small frequency drifts. \han{note}
A typical observation involves $1 \times 10^{10}$ wave cycles, and a coherent search must track the phase to better than half a cycle over the full observation. 
Here we consider the audio analog of a tone that wanders in frequency.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.48\textwidth]{figures/viterbiDiagramSizes.pdf}
\caption{\label{fig:viterbi}
Schematic diagram of the Viterbi algorithm. 
The circular nodes represent elements in a time-frequency grid, labeled $t_0$ to $t_{N_t}$ (left to right) in time and $f_0$ to $f_{N_f}$ (bottom to top) in frequency. 
The size of the nodes represents the likelihood of a signal being present in each time-frequency element. 
The small to large sizing corresponds to low to high likelihood values in arbitary units for this schematic. 
The color gradient (online only) also represents the likelihood values where the dark blue corresponds to low likelihood and light green corresponds to high likelihood. 
The objective is to find the most probable path of a signal through the grid from left to right.
All possible paths are shown by the lines. 
The black lines show the best path to each node and the gray lines show rejected paths. 
Some routes through the grid result in dead-ends with respect to optimally reaching the other side, such as the path ending at $t_1$ marked (a).
At $t=t_{N_t}$ the algorithm chooses the terminating frequency node which has the highest value given by Eqn.~\ref{eqn:pathprobability}, marked (b). 
The Viterbi path is the path leading to this node, highlighted by the thick-black lines, from (b) to the start at (c). 
}
\end{center}
\end{figure}

@@

In continuous gravitational wave searches, the signal amplitude is expected to be small in comparison to the noise and its frequency can change unpredictably over time.
The Viterbi algorithm's strength lies in its ability to track such signals through the data even in the presence of comparatively loud noise fluctuations in the time-frequency bins.
In the following section, we present the results of using the Viterbi algorithm with the table-top interferometer data.

>>>>

A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars, or variable accretion from a stellar companion for neutron stars in binaries such as LMXBs (see Ref.~\cite{SuvorovaEtAl:2017} and references therein).
Although continuous gravitational waves are nearly monochromatic, the long observing times ($\sim 1\,{\rm yr}$) mean that searches can be impacted by very small frequency drifts. 
A typical observation involves $1 \times 10^{10}$ wave cycles, and a coherent search must track the phase to better than half a cycle over the full observation. 
Here we consider the audio analog of a tone that wanders in frequency.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.48\textwidth]{figures/viterbiDiagramSizes.pdf}
\caption{\label{fig:viterbi}
Schematic diagram of the Viterbi algorithm. 
The circular nodes represent elements in a time-frequency grid, labeled $t_0$ to $t_{N_t}$ (left to right) in time and $f_0$ to $f_{N_f}$ (bottom to top) in frequency. 
The size of the nodes represents the likelihood of a signal being present in each time-frequency element. 
The small to large sizing corresponds to low to high likelihood values in arbitrary units for this schematic. 
The color gradient (online only) also represents the likelihood values where the dark blue corresponds to low likelihood and light green corresponds to high likelihood. 
The objective is to find the most probable path of a signal through the grid from left to right.
All possible paths are shown by the lines. 
The black lines show the best path to each node and the gray lines show rejected paths. 
Some routes through the grid result in dead-ends with respect to optimally reaching the other side, such as the path ending at $t_1$ marked (a).
At $t=t_{N_t}$ the algorithm chooses the terminating frequency node which has the highest value given by Eqn.~\ref{eqn:pathprobability}, marked (b). 
The Viterbi path is the path leading to this node, highlighted by the thick-black lines, from (b) to the start at (c). 
}
\end{center}
\end{figure}

@@

In continuous gravitational wave searches, the signal amplitude is expected to be small in comparison to the noise and its frequency can change unpredictably over time. % how to hyphenate ``continuous gravitational wave searches''?
The Viterbi algorithm's strength lies in its ability to track such signals through the data even in the presence of comparatively loud noise fluctuations in the time-frequency bins.
In the following section, we present the results of using the Viterbi algorithm with the table-top interferometer data.
